---
title: "Exercise 8.x: The fish tank dataset - solution"   
author: "Lieven Clement and Jeroen Gilis"
date: "statOmics, Ghent University (https://statomics.github.io)"  
output:
    html_document:
      code_download: true    
      theme: cosmo
      toc: true
      toc_float: true
      highlight: tango
      number_sections: true
---

Import libraries

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

# Background 

Researchers want to assess the effect of three different diets on the weight 
gain of fish. They have set up an experiment with 12 different tanks of fish. 
Each tank contains the same number of fish. The weight of 6 fish in each tank 
was measured at the beginning and the end of the experiment. The researchers 
recorded the weight gain.

# Experimental design


# Data import

```{r}
fish <- read.table("https://raw.githubusercontent.com/statOmics/PSLS21/data/fishTank.txt",header=TRUE)
head(fish)
```

# Tidy data

```{r}

```

# Data exploration

Make an appropriate plot to explore the data. 

```{r}

```

# Analysis

Specify the model that you will use for answering the research question.


## Model assumptions




## Overall effect of diet

## Post-hoc analysis

## Conclusion

Formulate your conclusion in terms of the research question. 
